---
title: "Gov 50 Final Project"
author: "Thomas Batties"
description: "My final project"
output:
  distill::distill_article:
    self_contained: false
---


## Project thoughts

I am interested in researching how, in the NBA, minutes per game affects offensive rebounds per game.


Does a player's minutes per game affect their RAPTOR +/-? In this study, I will examine the possible correlation between these two statistics. For context, RAPTOR +/- is a basketball-exclusive stat that measures by how many points a player boosts or hinders their team's performance. My hypothesis is that  a player who receives more minutes per game will have a higher RAPTOR +/-. Generally, better players will play more in a game, and they will have more of a positive effect on the game then their teammates who play less. Since a player's opportunity to record RAPTOR +/- is heavily dependent upon how many minutes they get, minutes is the explanatory variable. The outcome variable is RAPTOR +/-. The data would reflect my hypothesis if it showed a direct positive correlation between minutes per game and RAPTOR +/-.

```{r}
library(tidyverse)
library(ggplot2)
nba <- read_csv("nba-data-historical.csv")
nba
```


```{r}
plot1 <- nba |> 
  group_by(age) |>
  ggplot(data = nba,
         mapping = aes(x = `Raptor+/-`,
                       y = MPG)) + geom_point(alpha = 0.5)

plot1
```
